Titeln på avhandlingens språk

### 2022-12-19 abstract.tex
glossaries

\begin{abstract}
    \ac{WCA} applications present a challenge to benchmark and characterize due to their human-in-the-loop nature.
    Employing user testing to optimize system parameters is generally not feasible, given the scope of the problem and the number of observations needed to detect small but important effects in controlled experiments.
    Considering the intended mass-scale deployment of \ac{WCA} applications in the future, there exists a need for tools enabling human-independent benchmarking.

    We present in this paper the first model for the complete end-to-end emulation of humans in \ac{WCA}.
    We build this model through statistical analysis of data collected from previous work in this field, and demonstrate its utility by studying application task durations.
    Compared to first-order approximations, our model shows a \textasciitilde\SI{36}{\percent} larger gap between step execution times at high system impairment versus low.
    We further introduce a novel framework for stochastic optimization of resource consumption-responsiveness tradeoffs in \ac{WCA}, and show that by combining this framework with our realistic model of human behavior, significant reductions of up to \SI{50}{\percent} in number processed frame samples and \SI{20}{\percent} in energy consumption can be achieved with respect to the state-of-the-art.
\end{abstract}
\acresetall%

>>>>

\begin{abstract}
    \gls{WCA} applications present a challenge to benchmark and characterize due to their human-in-the-loop nature.
    Employing user testing to optimize system parameters is generally not feasible, given the scope of the problem and the number of observations needed to detect small but important effects in controlled experiments.
    Considering the intended mass-scale deployment of \gls{WCA} applications in the future, there exists a need for tools enabling human-independent benchmarking.

    We present in this paper the first model for the complete end-to-end emulation of humans in \gls{WCA}.
    We build this model through statistical analysis of data collected from previous work in this field, and demonstrate its utility by studying application task durations.
    Compared to first-order approximations, our model shows a \textasciitilde\SI{36}{\percent} larger gap between step execution times at high system impairment versus low.
    We further introduce a novel framework for stochastic optimization of resource consumption-responsiveness tradeoffs in \gls{WCA}, and show that by combining this framework with our realistic model of human behavior, significant reductions of up to \SI{50}{\percent} in number processed frame samples and \SI{20}{\percent} in energy consumption can be achieved with respect to the state-of-the-art.
\end{abstract}
\acresetall%

### 2022-12-19 appendices.tex
Revert "fixes for IEEE ToMC"
This reverts commit 18f0732eab5bbfdf6d172046f0725d1b21ce31c6.
glossaries

\clearpage%
\pagestyle{empty}%
\appendix{}\label[appendixwithoutnumber]{appx1}
% \subsection*{Sampling optimisation}

@@

To find 
% the expected wait 
$\mathbb{E}[\mathcal{W}]$, we use the conditional \ac{CDF} of the execution time.
\begin{multline*}
\mathbb{P}(\mathcal{W}=t_n-\mathcal{T}\leq t\,\big\vert\,t_{n-1}<\mathcal{T}\leq t_n)\\=\dfrac{\mathbb{P}(\mathcal{T}\geq t_n-t\,,\,t_{n-1}<\mathcal{T}\leq t_n)}{\mathbb{P}(t_{n-1}<\mathcal{T}\leq t_n)}.
\end{multline*}
The numerator is degenerate when $t\!<\!0$ or $t\!>\!(t_n\!-\!t_{n-1})$.
Thus, we are only interested in $0\!\leq\!t\!\leq\! (t_n-t_{n-1})$.
Let $F_{\mathcal{T}}$, $\bar{F}_{\mathcal{T}}$ and $f_{\mathcal{T}}$ correspond to the \ac{CDF}, \ac{CCDF} and \ac{PDF} of the execution time distribution.
\begin{alignat}{1}
\!\!\!\Rightarrow\mathbb{P}(\mathcal{W}\leq t\,\big\vert\,t_{n-1}<\mathcal{T}\leq t_n)&=\dfrac{\mathbb{P}(t_n-t\leq\mathcal{T}\leq t_n)}{F_\mathcal{T}(t_n)-F_\mathcal{T}(t_{n-1})}\nonumber\\
&\approx\dfrac{F_\mathcal{T}(t_n)-F_\mathcal{T}(t_{n}-t)}{F_\mathcal{T}(t_n)-F_\mathcal{T}(t_{n-1})}\label{Apx1}
\end{alignat}
Here, \cref{Apx1} is an approximation merely for mathematical maturity due to the slackness of the first inequality in the numerator.
We expand $F_\mathcal{T}(t_{n}-t)$ and $F_\mathcal{T}(t_{n-1})$ using Taylor series. Thus we can write the \ac{CCDF} as
\begin{multline*}
    =\Big(F_\mathcal{T}(t_n)-\\\big(F_\mathcal{T}(t_n)+f_\mathcal{T}(t_n)(-t)+f'_\mathcal{T}(t_n)(-t)^2/2!+\dots\big)\Big)\\
    \div \Big(F_\mathcal{T}(t_n)-\big(F_\mathcal{T}(t_n)+f_\mathcal{T}(t_n)(t_{n-1}-t_n)\\+f'_\mathcal{T}(t_n)(t_{n-1}-t_n)^2/2!+\dots\big)\Big).
\end{multline*}
Simplifying and approximating by ignoring the higher order terms, we arrive at
\begin{alignat}{1}
% &=\dfrac{F_\mathcal{T}(t_n)-\big(F_\mathcal{T}(t_n)+f_\mathcal{T}(t_n)(-t)+f'_\mathcal{T}(t_n)(-t)^2/2!+\dots\big)}{F_\mathcal{T}(t_n)-\big(F_\mathcal{T}(t_n)+f_\mathcal{T}(t_n)(t_{n-1}-t_n)+f'_\mathcal{T}(t_n)(t_{n-1}-t_n)^2/2!+\dots\big)}\\&
\mathbb{P}(\mathcal{W}\leq t\,\big\vert\,t_{n-1}<\mathcal{T}\leq t_n)&\approx\dfrac{tf_\mathcal{T}(t_n)}{(t_{n}-t_{n-1})f_\mathcal{T}(t_n)}\label{Apx2}\\
\Rightarrow\mathbb{P}(\mathcal{W}> t\,|\,t_{n-1}<\mathcal{T}\leq t_n)&= 1-\dfrac{t}{(t_{n}-t_{n-1})}\nonumber
\end{alignat}

Next, using the above \ac{CCDF}, we find the conditional expectation of $\mathcal{W}$.
\begin{alignat*}{1}
% \mathbb{P}(\mathcal{W}> t\,|\,t_{n-1}<\mathcal{T}\leq t_n)&\approx 1-\dfrac{t}{(t_{n}-t_{n-1})}\\
\Rightarrow \mathbb{E}[\mathcal{W}\,|\,t_{n-1}<\mathcal{T}\leq t_n]&=\smashoperator[r]{\int_{0}^{t_n-t_{n-1}}}\Big(1-\dfrac{t}{(t_{n}-t_{n-1})}\Big)\,\mathrm{d}t\\
&=\dfrac{(t_n-t_{n-1})}{2}.
\end{alignat*}
% However, we can also approximate $(t_n-t_{n-1})$, the sampling interval to the inverse of the instantaneous sampling frequency $r(t)$. This approximation (\ref{Apx3}) is valid as long as $r(t)$ is varying slowly between two consecutive sampling instants. That is,
If $r(t)$ is varying slowly between two consecutive sampling instants due to the closeness of two sampling intervals, we can approximate the sampling interval $(t_n\!-\!t_{n-1})$ as
\begin{alignat}{1}
 (t_n-t_{n-1})&\approx\dfrac{1}{r(t)},\;\forall t,n:t_{n-1}\!<\!t\!\leq\!t_n.\label{Apx3}\\
% \end{alignat}
% As a result, we can compute the expected wait as,
% \begin{alignat}{1}
\Rightarrow \mathbb{E}[\mathcal{W}]&=\int_{0}^{\infty}\mathbb{E}[\mathcal{W}\,|\,t_{n-1}<\mathcal{T}\leq t_n]f_\mathcal{T}(t)\,\mathrm{d}t\nonumber\\
&=\int_{0}^{\infty}\dfrac{1}{2r(t)}f_\mathcal{T}(t)\,\mathrm{d}t.\label{Ew}
\end{alignat}
We can thus find the energy penalty using \cref{Es} and \cref{Ew} as
\begin{alignat*}{1}
\mathcal{E}&=\int_{0}^{\infty}\Big(\alpha\int_{x=0}^{t}r(x)\,\mathrm{d}x+\beta\dfrac{1}{2r(t)}\Big)f_{\mathcal{T}}(t)\,\mathrm{d}t.\label{epsilon_eulerForm}\\
\intertext{%
    Let $g(t)=\int_{0}^{t}r(x)\dif x$.
    Then  $g'(t)=\tfrac{\mathrm{d}}{\mathrm{d}t}g(t)=r(t)$.
    That is,
}
\mathcal{E}&=\int_{0}^{\infty}\Big(\alpha g(t)+\dfrac{\beta}{2g'(t)}\Big)f_{\mathcal{T}}(t)\,\mathrm{d}t
\end{alignat*}

As per the Euler-Lagrange equation from the calculus of variations~\cite{Bellman1954Dynamic,Arfken2015Calculus}, the extreme value of $\mathcal{E}$ is obtained at 
\begin{alignat}{1}
r^*(t)&=\sqrt{\dfrac{\beta f_\mathcal{T}(t)}{2\alpha\bar{F}_\mathcal{T}(t)}}.\\
\intertext{Thus, for a Rayleigh distributed $\mathcal{T}$ with parameter $\sigma$,}
r^*(t)&=\sqrt{\dfrac{\beta t}{2\alpha\sigma^2}}\\
\Rightarrow &\int_{t_n}^{t_{n+1}}\!\!\!\sqrt{\dfrac{\beta t}{2\alpha\sigma^2}}\,\mathrm{d}t=1,\;\forall n\geq1\nonumber\tag{from \eqref{rt}}\\
\Rightarrow &\;t_{n+1}^{\frac{3}{2}}-t_{n}^{\frac{3}{2}}=3\sigma\!\sqrt{\tfrac{\alpha}{2\beta}}\nonumber\\
\intertext{We have $t_0=0$. Substituting $n=1,2,\dots$, in order, in the above equation provides us our final result}
&\;t_n=\Big(3\sigma\!\sqrt{\tfrac{\alpha}{2\beta}}\Big)^{\frac{2}{3}}n^{\frac{2}{3}}\label{tnRayleigh}.
x\end{alignat}
% Note that throughout the paper we use a exGaussian distribution but not Rayleigh. This is because, 
Note that, due to the close similarity in their density functions, the task times can be approximated fairly equally to an \acl{exGaussian} distribution as well as a Rayleigh distribution, with the former attracting more attention from works like~\cite{Rohrer1994analysis,Palmer2011shapes,Marmolejo2022generalised}.
This is the reason why the above results are applicable in this work where we have predominantly considered \ac{exGaussian} distribution.
Furthermore, we have also verified the closeness of the results as well as the validity of the approximations made in the proofs using distribution fitting and simulations.

>>>>

\appendix{}\label[appendixwithoutnumber]{appx1}
% \subsection*{Sampling optimisation}

@@

To find 
% the expected wait 
$\mathbb{E}[\mathcal{W}]$, we use the conditional \gls{CDF} of the execution time.
\begin{multline*}
\mathbb{P}(\mathcal{W}=t_n-\mathcal{T}\leq t\,\big\vert\,t_{n-1}<\mathcal{T}\leq t_n)\\=\dfrac{\mathbb{P}(\mathcal{T}\geq t_n-t\,,\,t_{n-1}<\mathcal{T}\leq t_n)}{\mathbb{P}(t_{n-1}<\mathcal{T}\leq t_n)}.
\end{multline*}
The numerator is degenerate when $t\!<\!0$ or $t\!>\!(t_n\!-\!t_{n-1})$.
Thus, we are only interested in $0\!\leq\!t\!\leq\! (t_n-t_{n-1})$.
Let $F_{\mathcal{T}}$, $\bar{F}_{\mathcal{T}}$ and $f_{\mathcal{T}}$ correspond to the \gls{CDF}, \gls{CCDF} and \gls{PDF} of the execution time distribution.
\begin{alignat}{1}
\!\!\!\Rightarrow\mathbb{P}(\mathcal{W}\leq t\,\big\vert\,t_{n-1}<\mathcal{T}\leq t_n)&=\dfrac{\mathbb{P}(t_n-t\leq\mathcal{T}\leq t_n)}{F_\mathcal{T}(t_n)-F_\mathcal{T}(t_{n-1})}\nonumber\\
&\approx\dfrac{F_\mathcal{T}(t_n)-F_\mathcal{T}(t_{n}-t)}{F_\mathcal{T}(t_n)-F_\mathcal{T}(t_{n-1})}\label{Apx1}
\end{alignat}
Here, \cref{Apx1} is an approximation merely for mathematical maturity due to the slackness of the first inequality in the numerator.
We expand $F_\mathcal{T}(t_{n}-t)$ and $F_\mathcal{T}(t_{n-1})$ using Taylor series. Thus we can write the \gls{CCDF} as
\begin{multline*}
    =\Big(F_\mathcal{T}(t_n)-\\\big(F_\mathcal{T}(t_n)+f_\mathcal{T}(t_n)(-t)+f'_\mathcal{T}(t_n)(-t)^2/2!+\dots\big)\Big)\\
    \div \Big(F_\mathcal{T}(t_n)-\big(F_\mathcal{T}(t_n)+f_\mathcal{T}(t_n)(t_{n-1}-t_n)\\+f'_\mathcal{T}(t_n)(t_{n-1}-t_n)^2/2!+\dots\big)\Big).
\end{multline*}
Simplifying and approximating by ignoring the higher order terms, we arrive at
\begin{alignat}{1}
% &=\dfrac{F_\mathcal{T}(t_n)-\big(F_\mathcal{T}(t_n)+f_\mathcal{T}(t_n)(-t)+f'_\mathcal{T}(t_n)(-t)^2/2!+\dots\big)}{F_\mathcal{T}(t_n)-\big(F_\mathcal{T}(t_n)+f_\mathcal{T}(t_n)(t_{n-1}-t_n)+f'_\mathcal{T}(t_n)(t_{n-1}-t_n)^2/2!+\dots\big)}\\&
\mathbb{P}(\mathcal{W}\leq t\,\big\vert\,t_{n-1}<\mathcal{T}\leq t_n)&\approx\dfrac{tf_\mathcal{T}(t_n)}{(t_{n}-t_{n-1})f_\mathcal{T}(t_n)}\label{Apx2}\\
\Rightarrow\mathbb{P}(\mathcal{W}> t\,|\,t_{n-1}<\mathcal{T}\leq t_n)&= 1-\dfrac{t}{(t_{n}-t_{n-1})}\nonumber
\end{alignat}

Next, using the above \gls{CCDF}, we find the conditional expectation of $\mathcal{W}$.
\begin{alignat*}{1}
% \mathbb{P}(\mathcal{W}> t\,|\,t_{n-1}<\mathcal{T}\leq t_n)&\approx 1-\dfrac{t}{(t_{n}-t_{n-1})}\\
\Rightarrow \mathbb{E}[\mathcal{W}\,|\,t_{n-1}<\mathcal{T}\leq t_n]&=\smashoperator[r]{\int_{0}^{t_n-t_{n-1}}}\Big(1-\dfrac{t}{(t_{n}-t_{n-1})}\Big)\,\mathrm{d}t\\
&=\dfrac{(t_n-t_{n-1})}{2}.
\end{alignat*}
% However, we can also approximate $(t_n-t_{n-1})$, the sampling interval to the inverse of the instantaneous sampling frequency $r(t)$. This approximation (\ref{Apx3}) is valid as long as $r(t)$ is varying slowly between two consecutive sampling instants. That is,
If $r(t)$ is varying slowly between two consecutive sampling instants due to the closeness of two sampling intervals, we can approximate the sampling interval $(t_n\!-\!t_{n-1})$ as
\begin{alignat}{1}
 (t_n-t_{n-1})&\approx\dfrac{1}{r(t)},\;\forall t,n:t_{n-1}\!<\!t\!\leq\!t_n.\label{Apx3}\\
% \end{alignat}
% As a result, we can compute the expected wait as,
% \begin{alignat}{1}
\Rightarrow \mathbb{E}[\mathcal{W}]&=\int_{0}^{\infty}\mathbb{E}[\mathcal{W}\,|\,t_{n-1}<\mathcal{T}\leq t_n]f_\mathcal{T}(t)\,\mathrm{d}t\nonumber\\
&=\int_{0}^{\infty}\dfrac{1}{2r(t)}f_\mathcal{T}(t)\,\mathrm{d}t.\label{Ew}
\end{alignat}
We can thus find the energy penalty using \cref{Es} and \cref{Ew} as
\begin{alignat*}{1}
\mathcal{E}&=\int_{0}^{\infty}\Big(\alpha\int_{x=0}^{t}r(x)\,\mathrm{d}x+\beta\dfrac{1}{2r(t)}\Big)f_{\mathcal{T}}(t)\,\mathrm{d}t.\label{epsilon_eulerForm}\\
\intertext{%
    Let $g(t)=\int_{0}^{t}r(x)\dif x$.
    Then  $g'(t)=\tfrac{\mathrm{d}}{\mathrm{d}t}g(t)=r(t)$.
    That is,
}
\mathcal{E}&=\int_{0}^{\infty}\Big(\alpha g(t)+\dfrac{\beta}{2g'(t)}\Big)f_{\mathcal{T}}(t)\,\mathrm{d}t
\end{alignat*}

As per the Euler-Lagrange equation from the calculus of variations~\cite{Bellman1954Dynamic,Arfken2015Calculus}, the extreme value of $\mathcal{E}$ is obtained at 
\begin{alignat}{1}
r^*(t)&=\sqrt{\dfrac{\beta f_\mathcal{T}(t)}{2\alpha\bar{F}_\mathcal{T}(t)}}.\\
\intertext{Thus, for a Rayleigh distributed $\mathcal{T}$ with parameter $\sigma$,}
r^*(t)&=\sqrt{\dfrac{\beta t}{2\alpha\sigma^2}}\\
\Rightarrow &\int_{t_n}^{t_{n+1}}\!\!\!\sqrt{\dfrac{\beta t}{2\alpha\sigma^2}}\,\mathrm{d}t=1,\;\forall n\geq1\nonumber\tag{from \eqref{rt}}\\
\Rightarrow &\;t_{n+1}^{\frac{3}{2}}-t_{n}^{\frac{3}{2}}=3\sigma\!\sqrt{\tfrac{\alpha}{2\beta}}\nonumber\\
\intertext{We have $t_0=0$. Substituting $n=1,2,\dots$, in order, in the above equation provides us our final result}
&\;t_n=\Big(3\sigma\!\sqrt{\tfrac{\alpha}{2\beta}}\Big)^{\frac{2}{3}}n^{\frac{2}{3}}\label{tnRayleigh}.
x\end{alignat}
% Note that throughout the paper we use a exGaussian distribution but not Rayleigh. This is because, 
Note that, due to the close similarity in their density functions, the task times can be approximated fairly equally to an \acl{exGaussian} distribution as well as a Rayleigh distribution, with the former attracting more attention from works like~\cite{Rohrer1994analysis,Palmer2011shapes,Marmolejo2022generalised}.
This is the reason why the above results are applicable in this work where we have predominantly considered \gls{exGaussian} distribution.
Furthermore, we have also verified the closeness of the results as well as the validity of the approximations made in the proofs using distribution fitting and simulations.
